---
title: "simpleRegress"
author: "Chris Gonzalez "
date: "Summer 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
The plot chart illustrates a bear with a neck measuring 20 inches, and it is anticipated that the bear's total length will reach approximately 57 inches. The residual standard error is 5.376 and the R- squared is .7524.

```{r}
mychest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(mychest)
summary.lm(mychest)
```
The residual standard error is 4.939, indicating a lower value in comparison to the initial length and neck measurements. This suggests that the current model provides a more accurate basis for comparison. Moreover, the chest measurement is ranked higher than the neck measurement.


```{r}
myhead <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(myhead)
summary.lm(myhead)
```
The residual standard error of 7.32 of length compared to the head width is a higher value when compared to the other two measurements. As a result, it would be ranked in third place, indicating that it is the least favorable approach for analysis.

```{r}
myweight <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(myweight)
summary.lm(myweight)
```
The comparison between the length and weight yields a residual standard error of 5.432. Additionally, the adjusted R-squared value is calculated to be 0.7423. According to the current ranking system, this combination would be ranked in fourth place.

```{r}
myheadlen <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(myheadlen)
summary.lm(myheadlen)
```
The comparison between the length and head length reveals a residual standard error of 4.235, accompanied by an adjusted R-squared value of 0.8434. According to the current ranking system, the head length surpasses other measurements, securing the first position in terms of analysis.

```{r}
myage <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(myage)
summary.lm(myage)
```
The comparison between length and age results in a residual standard error of 7.511, accompanied by an adjusted R-squared value of 0.5073. Based on the current ranking system, the length and age combination is deemed the least effective for comparison and analysis, thus placing it in the last position. 

##  Conclusion 
 
## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

