---
title: "simpleRegress"
author: "Chris Gonzalez "
date: "Summer 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
The plot chart illustrates a bear with a neck measuring 20 inches, and it is anticipated that the bear's total length will reach approximately 57 inches. The residual standard error is 5.376 and the R- squared is .7524.

```{r}
mychest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(mychest)
summary.lm(mychest)
```
The residual standard error is 4.939, indicating a lower value in comparison to the initial length and neck measurements. This suggests that the current model provides a more accurate basis for comparison. Moreover, the chest measurement is ranked higher than the neck measurement.


```{r}
myhead <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(myhead)
summary.lm(myhead)
```
The residual standard error of 7.32 is a higher value when compared to the other two measurements. As a result, it would be ranked in third place, indicating that it is the least favorable approach for analysis.

```{r}
myweight <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(myweight)
summary.lm(myweight)
```



## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

